---
title: "individual2"
author: "Blake Inderski, btski"
date: "March 22, 2019"
output: html_document
---

```{r}
library("dplyr")
library("tidyr")
library("tidyverse")
library("ggplot2")
library("readxl")
```

```{r}
#load data
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

```{r}
defClean <- defense
offClean <- offense
bioClean <- biography

defClean[, "Name"] <- lapply(defClean[, "Name"], as.factor)
defClean[, "Opponent_Opponent"] <- lapply(defClean[, "Opponent_Opponent"], as.factor)
offClean[, "Name"] <- lapply(offClean[, "Name"], as.factor)
offClean[, "Opponent_Opponent"] <- lapply(offClean[, "Opponent_Opponent"], as.factor)
bioClean[, "Name"] <- lapply(bioClean[, "Name"], as.factor)

defClean <- mutate_if(defClean, is.character, as.numeric)
offClean <- mutate_if(offClean, is.character, as.numeric)
bioClean[, "Weight"] <- lapply(bioClean[, "Weight"], as.numeric)

bioClean <- bioClean %>% separate(Height, into=c("Feet", "Inch"), sep="-") %>% mutate(Height = as.numeric(Feet)*12+as.numeric(Inch))
bioClean <- bioClean[, c(1, 2, 9, 5, 6, 7, 8)]

str(defClean)
str(offClean)
str(bioClean)
```

```{r}
defClean <- defClean %>% gather(key = Statistic, value = Count, Tackles_Solo:Pass_PB)
plot1 <- ggplot(defClean, aes(x=Count)) + geom_histogram() + facet_wrap(~Statistic)
plot1
```
<br />The ability for the defense to directly force a turnover (Turnover_FF, Turnover_FR, Turnover_INT) is least common skill among the provided defensive characteristics.

```{r}
defCleanFilter <- defClean %>% filter(Opponent_Opponent %in% c("West Virginia", "Kansas State")) %>% filter(Statistic == "Tackles_Solo") %>% spread(key = Statistic, value = Count) %>% spread(key = Opponent_Opponent, value = Tackles_Solo)
#defCleanFilter[is.na(defCleanFilter)] <- 0 #replace null values with zero
plot2 <- ggplot(defCleanFilter, aes(x=`West Virginia`, y=`Kansas State`, na.rm = TRUE)) + 
  geom_jitter(width=0.1, height=0.1) + xlim(0,10) + ylim(0,10) + geom_abline(slope=1, intercept=0)
plot2
```
<br />A greater number of points lie above the abline (x=y), indicating that the defense played better in their game against Kansas State than against West Virginia.

```{r}
bioClean <- bioClean %>% separate(Hometown, into=c("City", "State"), sep=", ")
head(bioClean)
```

```{r}
stateCount <- bioClean %>% group_by(State) %>% summarize(count=n())
stateCount[order(-stateCount$count),]
```

```{r}
offBio <- inner_join(offClean, bioClean, by = "Name")
offHome <- offBio %>% select(Name, City, State) %>% unique()
offHome
```

```{r}
offReceive <- group_by(offBio, Name) %>% summarise(R_YDS_mean = mean(Receiving_YDS), R_YDS_std = sd(Receiving_YDS), Weight = mean(Weight))
offReceive
ggplot(offBio, aes(x = reorder(Name, Weight, FUN = mean), y=Receiving_YDS)) + geom_boxplot() + theme(axis.text.x=element_text(angle=90,hjust=1))
ggplot(offBio, aes(x=Weight, y=Receiving_YDS)) + geom_point()
```
<br />No obvious relationship exists between weight and receiving yards.

```{r}
defBio <- inner_join(defClean, bioClean, by = "Name")
defBio[, "Position"] <- lapply(defBio[, "Position"], as.factor)
defBio[, "Statistic"] <- lapply(defBio[, "Statistic"], as.factor)
defBio[, "Count"] <- lapply(defBio[, "Count"], as.numeric)
tacklesByPosition <- group_by(defBio, Position) %>% filter(Statistic == "Tackles_Solo") %>% summarise(tackles = sum(Count))
tacklesByPosition
```
Q: Which position had the most solo tackles?<br />
A: The defensive back position had the most solo tackles.